A self-consistency check for unitary propagation of Hawking quanta

### 2016-11-04 Draft_BH_shells.tex
grammar check

\begin{abstract}
The black hole information paradox presumes that quantum field theory in curved spacetime can provide unitary propagations from near-horizon modes to asymptotic Hawking quanta. 
Instead of invoking conjectural quantum-gravity effects to modify such assumption, we propose a self-consistency check.
We establish an analogy to Feynman's analysis of a double-slit experiment. 
Feynman showed that a unitary propagation of the interfering particles, namely ignoring the entanglement with the double-slit, becomes an arbitrarily reliable assumption when the screen to project the interference pattern goes to infinitely far away.
We argue for an analogous self-consistency check for quantum field theory in curved spacetime.
We apply it to the propagation of Hawking quanta and test whether ignoring the entanglement with the geometry also becomes arbitrarily reliable in the limit of a large black hole. 
We present curious results to suggest a negative answer, and we discuss how this loss of na\"ive unitarity in QFT might be related to a solution of the paradox based on the soft-hair-memory effect.
\end{abstract}

@@

The black hole information paradox \cite{Haw76a} was sharpened to show a self-inconsistency among the following three widely-believed statements \cite{AMPS}\footnote{Different physicists may believe in some of those more strongly than the others, but very few would have argued against any single statement without the explicit conflict in the information paradox.}:
\begin{itemize}
\item {\bf Unitary evaporation:} A black hole will totally evaporate; the formation and evaporation of a black hole is described by an asymptotic observer by a unitary $S$-matrix, whose size is the exponential of the black hole's Bekenstein entropy.
\item {\bf General relativity:} The collapse-Schwarzschild geometry given by general relativity is a valid description of the spacetime everywhere except for near the singularity.
Therefore, there is no drama while crossing the horizon.
\item {\bf Quantum field theory in curved spacetime:} Away from the singularity, we can apply local quantum field theory (QFT) which describes microscopic steps of evaporation. 
For large black holes, it manifests itself as unitary propagation of every near-horizon mode into every asymptotic Hawking quanta.
\end{itemize}
Basically, Statement Three demands that a near horizon mode is related to an asymptotic Hawking quanta by a unitary transformation, therefore carrying the same qubit of information. 
However, Statement One and Two demand this unique qubit to be maximally entangled with two distinct objects, violating monogamy. 
This conflict has inspired many different proposals to modify one of the above three statements. 
For example, information loss or remnant \cite{Bek94} modifies Statement One; 
firewall \cite{BraPir09,AMPS} or ER=EPR \cite{MalSus13} modifies Statement Two;
various proposals of nonlocal effects near the horizon \cite{Gid12,DodSil15,OsuPag16}, causal patch complementarity \cite{HuiYan13,IlgYan13,LowTho14}, or computational complementarity \cite{HH} modify Statement Three.\footnote{Various versions of complementarity basically claim that QFT breaks down if one applies it compare quantities which are in-principle not measurable (or difficult to measure) by the same observer.}

@@

Before presenting our argument, we will first explain its relation to some existing ideas.
The recent paper by Osuga and Page \cite{OsuPag16} provides an abstract framework of how modifying Statement Three resolves the paradox.
Basically, the near horizon mode is thermal as demanded by Statement Two.
But it propagates out in a non-unitary process, gaining the appropriate information to restore a unitary $S$-matrix as demanded by Statement One. 
For such abstract model to work in practice, this non-unitary propagation should be understood as interaction with a hidden system.
Here hidden simply means something that a na\"ive application of local QFT is ignorant to.
The obvious candidate of such hidden system is the geometry.

One proposal along this line of thoughts was discussed in \cite{NomVar12}. 
The idea was that the black hole acquires a large uncertainty during the evaporation process. 
If we treat the superposition of different classical geometries as a quantum system, it may qualify as the hidden system that is responsible for the non-unitary propagation of Hawking quanta. 
Unfortunately, the early version of such proposal can be defeated by counting entropy.
In order for such non-unitary process to restore the asymptotic $S$-matrix, the hidden system must store the required information, which is comparable to the amount of microscopic information carried by the black hole.
However, the black-hole-no-hair theorem states that two black holes with identical macroscopic parameters (mass, momentum, angular momentum, charge) are isometric to each other. 
This limits the number of different classical geometries to be parametrized by macroscopic parameters only.
As a result, the Hilbert space of superposed classical geometries has far fewer dimensions than what is necessary to carry the required amount of information.\footnote{We thank Raphael Bousso for private communication.}
Thus, even if a propagating Hawking quantum does interact with this hidden system, it cannot reproduce a unitary $S$-matrix; therefore such modification of Statement Three alone is insufficient to resolve the paradox.

The recent development on black-hole-soft-hairs \cite{HawPer16} hinted a way to revive this idea.
Basically, it was realized that two isometric (regions of) geometries should not always be considered as the same state.
If one maps the geometries with congruences of geodesics, two isometric regions with different maps are actually different states.
That is because changes in the mapping geodesics are measurable memory effects, and a consistent physical theory must be able to distinguish them.
This realization may dramatically increase the number of different states encoded in classical geometries.
A tentative counting of state on a black hole horizon yields exactly the required number such that a modification to Statement Three can indeed resolve the paradox \cite{HawPer16}.

In this paper, we will take another step that is complementary to the above progress. It is exciting to know that modifying Statement Three can in-principle resolve the paradox. 
However, it remains a great mystery of why should local QFT break down in low curvature regions. 
If we just demand that it breaks down in order to resolve the information paradox, then there is no reason why this solution is preferred over modifying Statement One or Two. 
We would like to argue that there is a self-consistency check that we should apply to QFT in curved spacetime, for a reason that is totally independent from the information paradox.
Applying such self-consistency check, we can show that a unitary propagation from a near horizon mode to an asymptotic Hawking quantum is already self-inconsistent. 
Thus it is only natural to modify Statement Three.

@@

The starting point of our argument involves no new physics at all.
We simply recognize that local QFT implies a standard assumption of {\bf subsystem unitarity}.
Quantum mechanical evolution of a full system is by-definition unitary. 
However in practice, we are not able to describe everything by quantum mechanics all together, so we describe only subsystems. 
Namely, we assume that the full system can be separated into a ``classical background'' and a ``quantum subsystem''. 
After such artificial separation, since we only use quantum mechanics to describe the quantum subsystem, it always appears to be unitary.
However, such unitarity should not be taken as a physical fact unless one specifically checks the nature of interactions with the background.

@@

In Sec.\ref{sec-BlackHole}, we apply the self-consistency check to the propagation of Hawking quanta in a close analogy to Feynman's analysis. 
The classical geometry plays the role of the double-slit, a background that can potentially entangle with the Hawking quanta wavefunction and ruin its subsystem unitarity. 
We then assume that the unknown quantum nature of geometry can be parametrize by a classical uncertainty with an unknown but fixed, gauge-invariant size. 
A classical wave plays the role of the interference pattern. 
When even a classical wave decoheres due to the uncertainty of geometry, there is no reason to believed that the quantized version of such wave has unitarity.
The common expectation is that in the limit of a large black hole, gravitational effects outside the horizon are arbitrarily weak, thus local QFT should be arbitrarily trustworthy.
By analogy to Feynman's analysis, we should expect that with a fixed uncertainty in the geometry, any effect on the classical wave should drop to zero when we take the large black hole limit.
Interestingly, what we found seems to be the opposite. 
We show that with a fixed geometric uncertainty, a classical wave suffers an {\it arbitrarily large} correction in the large black hole limit.
By analogy to Feynman's analysis, this is strong evidence that the apparent unitarity of QFT in this case is not a valid assumption.

@@

One might object that the Eq.~(\ref{eq-stdcls}) is not entirely classical, since the value of $\Delta y^{\rm ds}$ has to originate from the unknown quantum mechanics of the double-slit. 
This is a valid concern, and it actually shows another strength of Eq.~(\ref{eq-stdcls}). 
Indeed we do not know the value of $\Delta y^{\rm ds}$ just from classical physics, but it is very reasonable to assume that it is intrinsic to the double-slit. 
Namely, its value should be fixed if we move the screen further away. Eq.~(\ref{eq-stdcls}) shows that with a fixed $\Delta y^{\rm ds}$, we can always make $L$ large enough to satisfy this condition. 
Thus, at least at the level of gedanken experiment, one can always arrange a situation that the interference pattern is visible, thus the subsystem unitarity is valid.

@@

The double-slit interference is an example that both the classical coherence and quantum unitarity can be calculated explicitly and show that they follow the same condition.
Here, we advocate that such a relation is actually a general rule.
{\it Classical coherence, instead of thinking of it as a practicality of measurements, should be treated as a prerequisit to quantum unitarity.} 
Thus when classical coherence fails, not only have we no practical measurements to confirm the purity of the subsystem wavefunction, such wavefunctions {\it actually did not evolve unitarily.}
Whatever effect responsible for a classical uncertainty large enough to disrupt classical coherence must have also interacted with the subsystem quantum-mechanically and became entangled with it.

For quantum field theory in a curved spacetime, we always assume that the quantum field is a subsystem that never entangles with the geometry, thus remains unitary.
Since we do not know quantum gravity/geometry, we cannot directly check this assumption by an actual calculation of entanglement.
However, we can always check classical coherence, and by this rule we advocate, the answer directly proves/disproves subsystem unitarity.

@@

While such generalization to QFT seems straightforward in theory, there are a few subtleties in practice.
In the double-slit experiment, the position uncertainty of the double-slit was the only obvious uncertainty of the background that we need to keep track of.
It is also the reasonable thing to hold fixed while moving the projection screen away.
In QFT, the background is the entire geometry, and there are infinitely many ways a geometry can be uncertain.
For a full analogy to Feynman's analysis, not only do we need to parametrize those uncertainties, we also need to choose the appropriate combination to hold fixed as we take a similar limit.
At this moment, we do not have a general formalism to setup such self-consistency check that can be applied to any general geometry. 
In Sec.\ref{sec-BlackHole}, we propose a specific way to apply such self-consistency check to the propagation of Hawking quanta.
In addition to gaining insights into the information paradox, we hope to learn some lessons to inspire a more general formalism of such self-consistency check.

QFT theorists might think that a more direct self-consistency check is feasible.
Instead of parametrizing a classical geometric uncertainty, one can treat deviations from the background geometry as a field and quantize it as any other fields.
Then a standard QFT calculation of scattering with this new field (graviton) should capture all interactions with the geometry, and a small scattering amplitude should guarantee that the interaction is suppressed. 
The concern for such method is that IR issues for these scattering calculations cannot be unambiguously regulated like in global Minkowski space.
Since soft gravitons lead to real physical changes in the geometry that can be seen in memory effects \cite{Wei65,HeLys14}, an IR ambiguity signals our ignorance on how these geometric changes enter QFT.
Thus a direct graviton scattering calculation cannot be the full extent of a self-consistency check, and it is reasonable for our method to give extra constraints on the reliability of local QFT.

@@

The usual treatment of how a near-horizon mode becomes an asymptotic Hawking quantum is also following a fixed-background assumption.
The background is the Schwarzschild geometry.
\begin{equation}
ds^2 = -\left(1-\frac{2M}{r}\right)dt^2 + \frac{dr^2}{1-\frac{2M}{r}} + r^2d\Omega_2^2~.
\label{eq-metric}
\end{equation}
When one applies local quantum field theory to describe a field on this fixed background, the result is by-definition unitary, but that is not necessary a physical fact. 
Given how little we understand quantum gravity, a classical self-consistency check as we described in the previous section is a reasonable thing to do.

@@

\begin{itemize}
\item {\bf Classical coherence by tracking geodesics.} The classical mode of a massless field basically has its peaks and nodes following null geodesics.
Thus, instead of actually solving the mode function, we can perform a much simpler calculation to track geodesics.
We start with two out-going null geodesics with the separation of one period of such mode near the horizon, and then we calculate the change in their separations at spatial infinity.
As this change increases, one should gradually lose faith in the unitary propagation of such mode, since it becomes increasingly difficult to experimentally verify.
\item {\bf The background and the uncertainty.} The classical background in this case is the Schwarzschild geometry, which can be written as the metric $g_{\mu\nu}$ in Eq.~(\ref{eq-metric}). 
A classical uncertainty can be expressed as deviation from such metric, $\Delta g_{\mu\nu}$. 
There are two major challenges here.
\begin{enumerate}
\item There are many more variables than the double-slit experiment. 
We consider only one particular form of $\Delta g_{\mu\nu}$ in this paper---back-reactions from the presence of extra matter of zero total energy. 
It is simple to calculate and reflects the physical intuition of vacuum pair fluctuation.
\item The value of $\Delta g_{\mu\nu}$ is gauge dependent. 
By relating it to the presence of extra matter, we can express it as a gauge-invariant local quantity, which is the natural thing to hold fixed in our analysis.
\end{enumerate}
Note that if we were reporting a positive result, that such uncertainty upholds unitarity, then one should question the validity of learning a general lesson from a special example. However, we will show that this particular form of $\Delta g_{\mu\nu}$ is already sufficient to question unitarity, and we did not choose this form to specifically do that. Thus our limited analysis is sufficient to raise a reasonable doubt to the classical-background approximation.
\item {\bf Renormalization.} It is well-known that na\"ive applications of QFT often suffers from UV/IR divergences.
In our case, we will see that even in Minkowski space, geometric uncertainty already leads to a finite change between the two null geodesics.
We will simply ``renormalize'' that away by saying that if the black hole case leads to a similar change, it should not be taken as a serious problem.
Our surprising result is that in the black hole case, we get an infinitely larger change, which is difficult to blame on the usual divergence of QFT.
\end{itemize}

@@

In this paper, we proposed a self-consistency check of QFT in curved spacetime geometries.
We have shown that propagation from the near-horizon region of a large black hole seems to fail such check.
As explained in Sec.\ref{sec-BlackHole}, the consistency requires us to parametrize the uncertainty in the classical geometry.
We have chosen a particular form of uncertainty in our explicit calculation.
Our result is gauge invariant and makes physical sense, but there might be a more general way to parametrize such uncertainty to make our approach even more general.

In the hindsight, the physical meaning of our effect, Eq.~(\ref{eq-inv}), is quite clear.
The geometric uncertainty, parametrized by $(\sigma\Delta l)$, leads to a fixed physical change locally where it is.
When there is a large gravitational redshift, such effect is amplified.
In the large black hole limit, starting from a fixed physical distance from the horizon, the redshift gets arbitrarily large, so the effect gets amplified by an arbitrarily large factor.
It is natural to expect similar effects in other geometries with horizons or even just expansions.
We look forward to exploring those more general cases.

@@

While formulating the information paradox, it is customary to take the large black hole limit.
Such limit allows us to identify near-horizon modes with wavelength much shorter than the curvature scale, but much longer than the UV cut-off scale.
For these modes, we are confident that they start in local vacuum states of QFT as if they live in Minkowski space.
This large black hole limit, which is effectively a near-Minkowski limit of a near horizon region, is the key argument to eliminate many na\"ive solutions to the information paradox.

The main result in this paper shows that there is a competing effect which stops us from taking this convenient limit.
The geometric uncertainty at distance $l$ from the horizon will only affect near-horizon modes of wavelength shorter than $l$.
If we fix the wavelength of the relevant near-horizon modes and increase the black hole size, our result shows that the unitary propagation to asymptotic Hawking radiation becomes less and less trustworthy. 
Since the dimensionless ratio $(l/M)$ is the only relevant physical parameter that comes into the scaling both effects, there seems to be no middle-ground that can make both unitarity and local-Minkowski-ness arbitrarily reliable.
We think such competing nature is a good sign and will take us closer to the solution of information paradox.

@@

Next we introduce two shells of matter with equal and opposite ADM mass, $\Delta M$ and $-\Delta M$ where $|\Delta M| \ll M$, at distances $r_1 < r_2$ as shown in the top part of figure \ref{fig-setup}. 
We model these shells as infinitesimally thin and therefore use the Israel Junction Conditions (IJC) \cite{Isr66} with a delta function shell to analyze the effects of the shells. 
The metric remains the same as Eq.~(\ref{eq-metric}), except for the region between the two shells, which is given by
\begin{equation}
	ds^2 = - \left( 1 - \frac{2M'}{r} \right) d t'^2 + \left( 1 - \frac{2M'}{r} \right)^{-1} dr^2 + r^2 d \Omega_2^2~.	\label{2}
\end{equation}
Here $M' = M + \Delta M$. 
Due to spherical symmetry, we can demand that the angular and radial coordinates to be the same as the background metric. 
But the time coordinate will be different, so we denote that as $t'$.

@@

The time component $S^0_0$ is the surface energy density.
\begin{equation}
	\sigma = S^0_0 = \frac{1}{4 \pi r_1} \left( \left( 1 - \frac{2M}{r_1} \right)^\frac{1}{2} - \left( 1 - \frac{2(M+\Delta M)}{r_1} \right)^\frac{1}{2} \right).
\end{equation}
When $2\Delta M \ll (r_1-2M)$, this is approximately given by 
\begin{eqnarray}
	\sigma  =  \frac{1}{4 \pi r_1} \left[ \partial_X \left( 1 - \frac{2X}{r_1} \right)^\frac{1}{2} \right]_{X = M} \Delta M
	 =  \frac{\Delta M}{4 \pi r_1^2} g_{tt}^{-\frac{1}{2}}	\label{sig-Dm}
\end{eqnarray}
Next we need to replace the coordinate distance $\Delta r$ with the proper distance $\Delta l$ and the relation between them is given by
\begin{equation}
	\Delta l = \left( 1 - \frac{2M}{r_1} \right)^{-\frac{1}{2}} \Delta r = g_{tt}^{-\frac{1}{2}} \Delta r.	\label{Dl-Dr}
\end{equation}
These equations allow us to replacing $\Delta M$ by $\sigma$ and $\Delta r$ by $\Delta l$ in Eq.~(\ref{eq-result}) to get Eq.~(\ref{eq-inv}).

>>>>

\begin{abstract}
The black hole information paradox presumes that quantum field theory in curved spacetime can provide unitary propagations from near-horizon modes to asymptotic Hawking quanta. 
Instead of invoking conjectural quantum-gravity effects to modify such assumption, we propose a self-consistency check.
We establish an analogy to Feynman's analysis of a double-slit experiment. 
Feynman showed that a unitary propagation of the interfering particles, namely ignoring the entanglement with the double-slit, becomes an arbitrarily reliable assumption when the screen the interference pattern is projected onto goes to infinitely far away.
We argue for an analogous self-consistency check for quantum field theory in curved spacetime.
We apply it to the propagation of Hawking quanta and test whether ignoring the entanglement with the geometry also becomes arbitrarily reliable in the limit of a large black hole. 
We present curious results to suggest a negative answer, and we discuss how this loss of na\"ive unitarity in QFT might be related to a solution of the paradox based on the soft-hair-memory effect.
\end{abstract}

@@

The black hole information paradox \cite{Haw76a} was sharpened to show a self-inconsistency among the following three widely-believed statements \cite{AMPS}\footnote{Different physicists may believe in some of those more strongly than the others, but very few would have argued against any single statement without the explicit conflict in the information paradox.}:
\begin{itemize}
\item {\bf Unitary evaporation:} A black hole will totally evaporate; the formation and evaporation of a black hole is described by an asymptotic observer by a unitary $S$-matrix, whose size is the exponential of the black hole's Bekenstein entropy.
\item {\bf General relativity:} The collapse-Schwarzschild geometry given by general relativity is a valid description of the spacetime everywhere except for near the singularity.
Therefore, there is no drama while crossing the horizon.
\item {\bf Quantum field theory in curved spacetime:} Away from the singularity, we can apply local quantum field theory (QFT) which describes microscopic steps of evaporation. 
For large black holes, it manifests itself as unitary propagation of every near-horizon mode into every asymptotic Hawking quanta.
\end{itemize}
Basically, Statement Three demands that a near horizon mode is related to an asymptotic Hawking quanta by a unitary transformation, and therefore carries the same qubit of information. 
However, Statement One and Two demand this unique qubit to be maximally entangled with two distinct objects, violating monogamy. 
This conflict has inspired many different proposals to modify one of the above three statements. 
For example, information loss or remnant \cite{Bek94} modifies Statement One; 
firewall \cite{BraPir09,AMPS} or ER=EPR \cite{MalSus13} modifies Statement Two;
various proposals of nonlocal effects near the horizon \cite{Gid12,DodSil15,OsuPag16}, causal patch complementarity \cite{HuiYan13,IlgYan13,LowTho14}, or computational complementarity \cite{HH} modify Statement Three.\footnote{Various versions of complementarity basically claim that QFT breaks down if one applies it to compare quantities which are in-principle not measurable (or difficult to measure) by the same observer.}

@@

Before presenting our argument, we will first explain its relation to some existing ideas.
The recent paper by Osuga and Page \cite{OsuPag16} provides an abstract framework of how modifying Statement Three resolves the paradox.
Basically, the near horizon mode is thermal as demanded by Statement Two.
However, it propagates out in a non-unitary process, gaining the appropriate information to restore the unitary $S$-matrix demanded by Statement One. 
For this abstract model to work in practice, this non-unitary propagation should be understood as interaction with a hidden system.
Here hidden simply means something that a na\"ive application of local QFT is ignorant to.
The obvious candidate of such hidden system is the geometry.

One proposal along this line of thoughts was discussed in \cite{NomVar12}. 
The idea was that the black hole acquires a large uncertainty during the evaporation process. 
If we treat the superposition of different classical geometries as a quantum system, it may qualify as the hidden system that is responsible for the non-unitary propagation of Hawking quanta. 
Unfortunately, the early version of this proposal can be defeated by counting entropy.
In order for such non-unitary process to restore the asymptotic $S$-matrix, the hidden system must store the required information, which is comparable to the amount of microscopic information carried by the black hole.
However, the black-hole-no-hair theorem states that two black holes with identical macroscopic parameters (mass, momentum, angular momentum, charge) are isometric to each other. 
This limits the number of different classical geometries to be parametrized by macroscopic parameters only.
As a result, the Hilbert space of superposed classical geometries has far fewer dimensions than what is necessary to carry the required amount of information.\footnote{We thank Raphael Bousso for private communication.}
Thus, even if a propagating Hawking quantum does interact with this hidden system, it cannot reproduce a unitary $S$-matrix; therefore such modification of Statement Three alone is insufficient to resolve the paradox.

The recent development on black-hole-soft-hairs \cite{HawPer16} hinted a way to revive this idea.
Essentially, it was realized that two isometric (regions of) geometries should not always be considered as the same state.
If one maps the geometries with congruences of geodesics, two isometric regions with different maps are actually different states.
That is because changes in the mapping geodesics are measurable memory effects, and a consistent physical theory must be able to distinguish them.
This realization may dramatically increase the number of different states encoded in classical geometries.
A tentative counting of state on a black hole horizon yields exactly the required number such that a modification to Statement Three can indeed resolve the paradox \cite{HawPer16}.

In this paper, we will take another step that is complementary to the above progress. It is exciting to know that modifying Statement Three can, in-principle, resolve the paradox, yet it remains a great mystery why local QFT should break down in low curvature regions. 
If we just demand that it breaks down in order to resolve the information paradox, then there is no reason why this solution is preferred over modifying Statement One or Two. 
We would like to argue that there is a self-consistency check that we should apply to QFT in curved spacetime, for a reason that is totally independent from the information paradox.
Applying such self-consistency check, we can show that a unitary propagation from a near horizon mode to an asymptotic Hawking quantum is already self-inconsistent. 
Thus, it is only natural to modify Statement Three.

@@

The starting point of our argument involves no new physics at all.
We simply recognize that local QFT implies a standard assumption of {\bf subsystem unitarity}.
Quantum mechanical evolution of a full system is by-definition unitary. 
However in practice, we are not able to describe everything by quantum mechanics all together, so we describe only subsystems. 
Namely, we assume that the full system can be separated into a ``classical background'' and a ``quantum subsystem''. 
After this artificial separation, since we only use quantum mechanics to describe the quantum subsystem, it always appears to be unitary.
However, such unitarity should not be taken as a physical fact unless one specifically checks the nature of interactions with the background.

@@

In Sec.\ref{sec-BlackHole}, we apply the self-consistency check to the propagation of Hawking quanta in a close analogy to Feynman's analysis. 
The classical geometry plays the role of the double-slit, a background that can potentially entangle with the Hawking quanta wavefunction and ruin its subsystem unitarity. 
We then assume that the unknown quantum nature of geometry can be parametrize by a classical uncertainty with an unknown but fixed, gauge-invariant size. 
A classical wave plays the role of the interference pattern. 
When even a classical wave decoheres due to the uncertainty of geometry, there is no reason to believed that the quantized version of such wave has unitarity.
The common expectation is that in the limit of a large black hole, gravitational effects outside the horizon are arbitrarily weak, thus local QFT should be arbitrarily trustworthy.
By analogy to Feynman's analysis, we should expect that with a fixed uncertainty in the geometry, any effect on the classical wave should drop to zero when we take the large black hole limit.
Interestingly, what we found seems to be the opposite. 
We show that with a fixed geometric uncertainty, a classical wave suffers an {\it arbitrarily large} correction in the large black hole limit.
By analogy to Feynman's analysis, this is strong evidence that the apparent unitarity of QFT is, in this case, not a valid assumption.

@@

One might object that the Eq.~(\ref{eq-stdcls}) is not entirely classical, since the value of $\Delta y^{\rm ds}$ has to originate from the unknown quantum mechanics of the double-slit. 
This is a valid concern, and it actually shows another strength of Eq.~(\ref{eq-stdcls}). 
Indeed we do not know the value of $\Delta y^{\rm ds}$ just from classical physics, but it is very reasonable to assume that it is intrinsic to the double-slit. 
Namely, its value should be fixed if we move the screen further away. Eq.~(\ref{eq-stdcls}) shows that with a fixed $\Delta y^{\rm ds}$, we can always make $L$ large enough to satisfy this condition. 
Thus, at least at the level of gedanken experiment, one can always arrange a situation that the interference pattern is visible and subsystem unitarity is valid.

@@

The double-slit experiment is an example where both the classical coherence and quantum unitarity can be calculated explicitly and shown to follow the same condition.
Here, we advocate that such a relation is actually a general rule.
{\it Classical coherence, instead of thinking of it as a practicality of measurements, should be treated as a prerequisit to quantum unitarity.} 
Thus when classical coherence fails, not only have we no practical measurements to confirm the purity of the subsystem wavefunction, such wavefunctions {\it actually did not evolve unitarily.}
Whatever effect responsible for a classical uncertainty large enough to disrupt classical coherence must have also interacted with the subsystem quantum-mechanically and became entangled with it.

For quantum field theory in a curved spacetime, we always assume that the quantum field is a subsystem that never entangles with the geometry, and so remains unitary.
Since we do not know quantum gravity/geometry, we cannot directly check this assumption by an actual calculation of entanglement.
However, we can always check classical coherence and, by this rule we advocate, the answer directly proves/disproves subsystem unitarity.

@@

While such generalization to QFT seems straightforward in theory, there are a few subtleties in practice.
In the double-slit experiment, the position uncertainty of the double-slit was the only obvious uncertainty of the background that we need to keep track of.
It is also the reasonable thing to hold fixed while moving the projection screen away.
In QFT, the background is the entire geometry, and there are infinitely many ways a geometry can be uncertain.
For a full analogy to Feynman's analysis, not only do we need to parametrize those uncertainties, we also need to choose the appropriate combination to hold fixed as we take a similar limit.
At this moment, we do not have a general formalism to setup such self-consistency check that can be applied to any general geometry. 
In Sec.\ref{sec-BlackHole}, we propose a specific way to apply such a self-consistency check to the propagation of Hawking quanta.
In addition to gaining insights into the information paradox, we hope to learn some lessons to inspire a more general formalism of such self-consistency checks.

QFT theorists might think that a more direct self-consistency check is feasible.
Instead of parametrizing a classical geometric uncertainty, one can treat deviations from the background geometry as a field and quantize it as any other fields.
Then a standard QFT calculation of scattering with this new field (graviton) should capture all interactions with the geometry, and a small scattering amplitude should guarantee that the interaction is suppressed. 
The concern for this method is that IR issues for these scattering calculations cannot be unambiguously regulated as in global Minkowski space.
Since soft gravitons lead to real physical changes in the geometry that can be seen in memory effects \cite{Wei65,HeLys14}, an IR ambiguity signals our ignorance on how these geometric changes enter QFT.
Thus a direct graviton scattering calculation cannot be the full extent of a self-consistency check, and it is reasonable for our method to give extra constraints on the reliability of local QFT.

@@

The usual treatment of how a near-horizon mode becomes an asymptotic Hawking quantum also uses a fixed-background assumption.
The background is the Schwarzschild geometry.
\begin{equation}
ds^2 = -\left(1-\frac{2M}{r}\right)dt^2 + \frac{dr^2}{1-\frac{2M}{r}} + r^2d\Omega_2^2~.
\label{eq-metric}
\end{equation}
When one applies local quantum field theory to describe a field on this fixed background, the result is by-definition unitary, but that is not necessary a physical fact. 
Given how little we understand quantum gravity, a classical self-consistency check as we described in the previous section is a reasonable thing to do.

@@

\begin{itemize}
\item {\bf Classical coherence by tracking geodesics.} The classical mode of a massless field has its peaks and nodes following null geodesics.
Thus, instead of actually solving the mode function, we can perform a much simpler calculation to track geodesics.
We start with two out-going null geodesics with the separation of one period of such mode near the horizon, and then we calculate the change in their separations at spatial infinity.
As this change increases, one should gradually lose faith in the unitary propagation of such mode, since it becomes increasingly difficult to experimentally verify.
\item {\bf The background and the uncertainty.} The classical background in this case is the Schwarzschild geometry, which can be written as the metric $g_{\mu\nu}$ in Eq.~(\ref{eq-metric}). 
A classical uncertainty can be expressed as deviation from such metric, $\Delta g_{\mu\nu}$. 
There are two major challenges here:
\begin{enumerate}
\item There are many more variables than the double-slit experiment. 
We consider only one particular form of $\Delta g_{\mu\nu}$ in this paper---back-reactions from the presence of extra matter of zero total energy. 
It is simple to calculate and reflects the physical intuition of vacuum pair fluctuation.
\item The value of $\Delta g_{\mu\nu}$ is gauge dependent. 
By relating it to the presence of extra matter, we can express it as a gauge-invariant local quantity, which is the natural thing to hold fixed in our analysis.
\end{enumerate}
Note that if we were reporting a positive result, that such uncertainty upholds unitarity, then one should question the validity of learning a general lesson from a special example. However, we will show that this particular form of $\Delta g_{\mu\nu}$ is already sufficient to question unitarity, and we did not choose this form to specifically do that. Thus our limited analysis is sufficient to raise a reasonable doubt to the classical-background approximation.
\item {\bf Renormalization.} It is well-known that na\"ive applications of QFT often suffers from UV/IR divergences.
In our case, we will see that even in Minkowski space, geometric uncertainty already leads to a finite change between the two null geodesics.
We will simply ``renormalize'' that away by saying that if the black hole case leads to a similar change, it should not be taken as a serious problem.
Our surprising result is that in the black hole case, we get an infinitely larger change, which is difficult to blame on the usual divergence of QFT.
\end{itemize}

@@

In this paper, we proposed a self-consistency check of QFT in curved spacetime geometries.
We have shown that propagation from the near-horizon region of a large black hole seems to fail such check.
As explained in Sec.\ref{sec-BlackHole}, checking for consistency requires us to parametrize the uncertainty in the classical geometry.
We have chosen a particular form of uncertainty in our explicit calculation.
Our result is gauge invariant and makes physical sense, but there might be a more general way to parametrize such uncertainty to make our approach even more general.

In hindsight, the physical meaning of our effect, Eq.~(\ref{eq-inv}), is quite clear.
The geometric uncertainty, parametrized by $(\sigma\Delta l)$, leads to a fixed physical change locally.
When there is a large gravitational redshift, the effect is amplified.
In the large black hole limit, starting from a fixed physical distance from the horizon, the redshift gets arbitrarily large, so the effect gets amplified by an arbitrarily large factor.
It is natural to expect similar effects in other geometries with horizons or even just expansions.
We look forward to exploring those more general cases.

@@

While formulating the information paradox, it is customary to take the large black hole limit.
Such a limit allows us to identify near-horizon modes with wavelength much shorter than the curvature scale, but much longer than the UV cut-off scale.
For these modes, we are confident that they start in local vacuum states of QFT as if they live in Minkowski space.
This large black hole limit, which is effectively a near-Minkowski limit of a near horizon region, is the key argument to eliminate many na\"ive solutions to the information paradox.

The main result in this paper shows that there is a competing effect which stops us from taking this convenient limit.
The geometric uncertainty at distance $l$ from the horizon will only affect near-horizon modes of wavelength shorter than $l$.
If we fix the wavelength of the relevant near-horizon modes and increase the black hole size, our result shows that the unitary propagation to asymptotic Hawking radiation becomes less and less trustworthy. 
Since the dimensionless ratio $(l/M)$ is the only relevant physical parameter that comes into the scaling both effects, there seems to be no middle-ground that can make both unitarity and local-Minkowski-ness arbitrarily reliable.
We think such competition is a good sign and will take us closer to the solution of information paradox.

@@

Next we introduce two shells of matter with equal and opposite ADM mass, $\Delta M$ and $-\Delta M$ where $|\Delta M| \ll M$, at distances $r_1 < r_2$ as shown in the top part of figure \ref{fig-setup}. 
We model these shells as infinitesimally thin and therefore use the Israel Junction Conditions (IJC) \cite{Isr66} with a delta function shell to analyze the effects of the shells. 
The metric remains the same as Eq.~(\ref{eq-metric}), except for the region between the two shells, which is given by
\begin{equation}
	ds^2_{\rm region \ II} = - \left( 1 - \frac{2M'}{r} \right) d t'^2 + \left( 1 - \frac{2M'}{r} \right)^{-1} dr^2 + r^2 d \Omega_2^2~.	\label{2}
\end{equation}
Here $M' = M + \Delta M$. 
Due to spherical symmetry, we can demand that the angular and radial coordinates to be the same as the background metric. 
But the time coordinate will be different, so we denote that as $t'$.

@@

The time component $S^0_0$ is the surface energy density.
\begin{equation}
	\sigma = S^0_0 = \frac{1}{4 \pi r_1} \left( \left( 1 - \frac{2M}{r_1} \right)^\frac{1}{2} - \left( 1 - \frac{2(M+\Delta M)}{r_1} \right)^\frac{1}{2} \right).
\end{equation}
When $2\Delta M \ll (r_1-2M)$, this is approximately given by 
\begin{eqnarray}
	\sigma  =  \frac{-1}{4 \pi r_1} \left[ \partial_X \left( 1 - \frac{2X}{r_1} \right)^\frac{1}{2} \right]_{X = M} \Delta M
	 =  \frac{\Delta M}{4 \pi r_1^2} g_{tt}^{-\frac{1}{2}}	\label{sig-Dm}
\end{eqnarray}
Next we need to replace the coordinate distance $\Delta r$ with the proper distance $\Delta l$ and the relation between them is given by
\begin{equation}
	\Delta l = \left( 1 - \frac{2M}{r_1} \right)^{-\frac{1}{2}} \Delta r = g_{tt}^{-\frac{1}{2}} \Delta r.	\label{Dl-Dr}
\end{equation}
These equations allow us to replacing $\Delta M$ by $\sigma$ and $\Delta r$ by $\Delta l$ in Eq.~(\ref{eq-result}) to get Eq.~(\ref{eq-inv}).
